---
brand: 
  _brand.yml
execute: 
  echo: false
format: 
  typst:
    fig-width: 6
    fig-asp: 0.618
    title: "Labour Market Brief"
    subtitle: "{{< meta params.state >}}"
  html:
    title: "Labour Market Brief"
    subtitle: "{{< meta params.state >}}"
    output-ext: html
editor: visual
params:
  state: South Australia
---

```{r}
#| label: setup
#| include: false

library(dplyr)
library(stringr)
library(tidyr)
library(ggplot2)
library(lubridate)
library(reportabs)
library(scales)
library(showtext)

labour_force_briefing <- read_absdata("labour_force_briefing")
filter_list <- list(indicator = "Employed total", sex = "Persons", age = "Total (age)", state = params$state, series_type = "Seasonally Adjusted")

```

# Summary

The ABS Labour Force Survey data for **`r params$state`** released in `r release(labour_force_briefing, 'month', 1)` for `r release(labour_force_briefing, 'month')` showed that in **`r str_to_lower("Seasonally Adjusted")`** terms:

-   Total *employment* `r change(labour_force_briefing, filter_with = filter_list, ym = 'month')` people.
    -   Over the past year, employment in `r params$state` has `r growth(labour_force_briefing, filter_with = filter_list,  ym = 'year')`.
-   Total *unemployment* `r filter_list$indicator = "Unemployed total"; change(labour_force_briefing, filter_with = filter_list, ym = 'month')` people, and the *unemployment rate* `r filter_list$indicator = "Unemployment rate";  change(labour_force_briefing, filter_with = filter_list, ym = 'month')`.
    -   Over the past year, unemployment in `r params$state` has `r filter_list$indicator = "Unemployed total"; growth(labour_force_briefing, filter_with = filter_list, ym = 'year')`, and the unemployment rate has `r filter_list$indicator = "Unemployment rate"; growth(labour_force_briefing, filter_with = filter_list, ym = 'year')`.
-   Total *underutilisation* has `r filter_list$indicator = 'Underutilised total'; change(labour_force_briefing, filter_with = filter_list, ym = 'month')` people, and the *underutilisation rate* `r filter_list$indicator = "Underutilisation rate"` `r change(labour_force_briefing, filter_with = filter_list, ym = 'month')`.
    -   Over the past year, total underutilisation has `r filter_list$indicator = "Underutilised total"; growth(labour_force_briefing, filter_with = filter_list, ym = 'year')` and the underutilisation rate has `r filter_list$indicator = "Underutilisation rate"; growth(labour_force_briefing, filter_with = filter_list, ym = 'year')`.

**Nationally** in `r release(labour_force_briefing, 'month')`:

-   Total *employment* `r filter_list = list(indicator = "Employed total", sex = "Persons", age = "Total (age)", state = "Australia", series_type = "Seasonally Adjusted"); change(labour_force_briefing, filter_with = filter_list, ym = 'month')` people.
    -   Over the past year, employment in Australia has `r growth(labour_force_briefing, filter_with = filter_list, ym = 'year')`.
-   Total *unemployment* `r filter_list$indicator = "Unemployed total"; change(labour_force_briefing, filter_with = filter_list, ym = 'month')` people and the *unemployment rate* `r filter_list$indicator = "Unemployment rate"` `r change(labour_force_briefing,  filter_with = filter_list, ym = 'month')`
    -   Over the past year, unemployment in Australia has `r filter_list$indicator = "Unemployed total"; growth(labour_force_briefing, filter_with = filter_list,  ym = 'year')` and the *unemployment rate* has `r filter_list$indicator = "Unemployment rate"; growth(labour_force_briefing, filter_with = filter_list, ym = 'year')`.
-   Total *underutilisation* `r filter_list$indicator = 'Underutilised total'; change(labour_force_briefing, filter_with = filter_list, ym = 'month')` people, and the *underutilisation rate* `r filter_list$indicator = "Underutilisation rate"; change(labour_force_briefing, filter_with = filter_list,  ym = "month")`.
    -   Over the past year, total underutilisation has `r filter_list$indicator = "Underutilised total"` `r growth(labour_force_briefing, filter_with = filter_list,  ym = 'year')` and the underutilisation rate has `r filter_list$indicator = "Underutilisation rate"` `r growth(labour_force_briefing, filter_with = filter_list,  ym = 'year')`.

{{< pagebreak >}}

# Employment

`r {filter_list = list(indicator = "Employed total", state = params$state, sex = "Persons", age = "Total (age)", series_type = "Seasonally Adjusted")}`

-   In `r release(labour_force_briefing, 'month')`, `r str_to_lower("Seasonally Adjusted")` employment in `r params$state` `r change(labour_force_briefing, filter_with = filter_list,  ym = 'month')` people.
-   Compared with `r release(labour_force_briefing, 'month')` last year, total employment in `r params$state` has `r growth(labour_force_briefing, filter_with = filter_list,  ym = 'year')`.
-   Nationally, employment in `r params$state; filter_list$state = "Australia"; release(labour_force_briefing, 'month'); change(labour_force_briefing, filter_with = filter_list,  ym = 'month')` people.

```{r}
#| label: employment
#| out-width: "70%"
#| fig-align: center
filter_list$state = params$state
abs_plot(labour_force_briefing, over = filter_list)
```

## Full and Part Time Employment

`r {filter_list = list(state = params$state, indicator = "Employed full-time", sex = "Persons", series_type = "Seasonally Adjusted")}`

-   Full-time employment in `r params$state` `r change(labour_force_briefing,  filter_with = filter_list,  ym = 'month')` in `r release(labour_force_briefing, 'month')`, and has `r growth(labour_force_briefing, filter_with = filter_list,  ym = 'year')` over the year. `r {filter_list = list(indicator = "Employed part-time", series_type = "Seasonally Adjusted", state = params$state)}`
-   Part-time employment in `r params$state`, in original terms `r change(labour_force_briefing, filter_with = filter_list,  ym = 'month')` in `r release(labour_force_briefing, 'month')`. Part-time employment has `r growth(labour_force_briefing, filter_with = filter_list,  ym = 'year')` over the last year.

```{r}
#| label: full-part-time-employment
#| out-width: "70%"
#| fig-align: center
filter_list$indicator = c("Employed part-time", "Employed full-time")

abs_plot(labour_force_briefing, 
         over = filter_list, 
         compare_aus = FALSE) 

```

# Unemployment

`r {filter_list = list(indicator = "Unemployed total", state = params$state, series_type = "Seasonally Adjusted")}`

-   In `r release(labour_force_briefing, 'month')`, the total number of unemployed people in `r params$state` `r change(labour_force_briefing, filter_with = filter_list,  ym = 'month')`.

-   Compared with `r release(labour_force_briefing, 'month')` last year, unemployment in `r params$state` has `r growth(labour_force_briefing, filter_with = filter_list,  ym = 'year')` people. `r filter_list$state = "Australia"`

-   Nationally, total unemployment in `r release(labour_force_briefing, 'month')` `r change(labour_force_briefing, filter_with = filter_list,  ym = 'month')` people.

```{r}
#| label: unemployment-level
#| out-width: "70%"
#| fig-align: center
filter_list$state = params$state
abs_plot(labour_force_briefing, filter_list)
```

## Male and Female Unemployment Rates

`r {filter_list = list(sex = "Males", indicator = "Unemployment rate", series_type = "Seasonally Adjusted", state = params$state)}`

-   In `r release(labour_force_briefing, 'month')`, the `r str_to_lower("Seasonally Adjusted")` unemployment rate for males in `r params$state` `r change(labour_force_briefing, filter_with = filter_list,  ym = 'month')`.
-   Compared with `r release(labour_force_briefing, 'month')` last year male unemployment in `r params$state` has `r growth(labour_force_briefing, filter_with = filter_list,  ym = 'year')`.
-   In the same period, the `r str_to_lower("Seasonally Adjusted")` unemployment rate for females in `r params$state; filter_list$sex = "Females"; params$state; change(labour_force_briefing, filter_with = filter_list,  ym = 'month')`.
-   Compared with `r release(labour_force_briefing, 'month')` last year female unemployment in `r params$state` has `r growth(labour_force_briefing, filter_with = filter_list,  ym = 'year')`.

```{r}
#| label: male-female-unemployment
#| out-width: "70%"
#| fig-align: center
filter_list$sex = c("Males", "Females")
abs_plot(labour_force_briefing, over = filter_list, compare_aus = F)

```

# Underemployment

`r {filter_list = list(state = params$state, series_type = "Seasonally Adjusted", indicator = "Underemployment rate (proportion of labour force)")}`

-   The underemployment rate `r change(labour_force_briefing, filter_with = filter_list,  ym = 'month')` `r filter_list$indicator = "Underemployed total"`. The total number of people underemployed has `r change(labour_force_briefing, filter_with = filter_list, ym = 'month')` people.

-   Over the past year, total underemployment has `r growth(labour_force_briefing, filter_with = filter_list,  ym = 'year')`.

-   Nationally, the underemployment rate in `r release(labour_force_briefing,   ym = 'month')` `r filter_list$state = "Australia"; filter_list$indicator = "Underemployment rate (proportion of labour force)"` `r change(labour_force_briefing, filter_with = filter_list,  ym = 'month')` and the total number of people underemployed `r filter_list$indicator = "Underemployed total"` `r change(labour_force_briefing, filter_with = filter_list, ym = 'month')`.

```{r}
#| label: underemployment-rate
#| out-width: "70%"
#| fig-align: center
filter_list$state = params$state
filter_list$indicator = "Underemployment rate (proportion of labour force)"
abs_plot(labour_force_briefing, filter_list)
```

```{r}
#| label: underemployment-number
#| out-width: "70%"
#| fig-align: center
filter_list$state = params$state
filter_list$indicator = "Underemployed total"
abs_plot(labour_force_briefing, filter_list)

```

# Hours Worked

`r  {filter_list = list(state = params$state, series_type = "Seasonally Adjusted", indicator = "Monthly hours worked in all jobs")}`

-   The total number of hours worked in all jobs `r change(labour_force_briefing, filter_with = filter_list,  ym = 'month')`.

-   Over the past year, the number of hours worked in `r params$state` has `r growth(labour_force_briefing, filter_with = filter_list,  ym = 'year')`.

-   Nationally, `r filter_list$state = "Australia"` the total number of hours worked in `r release(labour_force_briefing, ym = 'month')` `r change(labour_force_briefing, filter_with = filter_list,  ym = 'month')`.

```{r}
#| label: hours-worked
#| out-width: "70%"
#| fig-align: center
filter_list$state = params$state
abs_plot(labour_force_briefing, filter_list)
```
