---
execute: 
  echo: false
format: 
  pdf:
    title: "Labour Market Brief"
    subtitle: "{{< meta params.state >}}"
    documentclass: scrartcl
    papersize: latter
    mainfont: Roboto
    fontsize: 10pt
    include-in-header: 
      - file: header.tex
editor: visual
params:
  series_type: Seasonally Adjusted
  state: South Australia
  years: 2017
---

```{r}
#| label: setup
#| include: false

library(dplyr)
library(stringr)
library(tidyr)
library(ggplot2)
library(lubridate)
library(reportabs)
library(scales)
library(ragg)

ragg_png <- function(..., res = 300) {
  agg_png(..., res = res, units = "in")
}
knitr::opts_chunk$set(dev = "ragg_png")

#Create underutilised total indicator

data <- list("labour_force" = read_absdata("labour_force") |>
                 tidyr::pivot_wider(id_cols = c(date, sex, state, series_type, unit, age),
                                    names_from = indicator,
                                    values_from = value) |>
                 dplyr::mutate(`Underutilised total` = `Underemployed total` + `Unemployed total`) |>
                 tidyr::pivot_longer(cols = "Employed total":"Underutilised total",
                                     names_to = "indicator",
                                     values_to = "value",
                                     values_drop_na = TRUE),
               "hours_worked" = read_absdata("hours_worked"))

labour_force <- data$labour_force
hours_worked <- data$hours_worked
filter_list <- list(indicator = "Employed total", sex = "Persons", age = "Total (age)", state = params$state, series_type = params$series_type)

```

\section{Summary}

The ABS Labour Force Survey data for **`r params$state`** released in `r release(labour_force, 'month', 1)` for `r release(labour_force, 'month')` showed that in **`r str_to_lower(params$series_type)`** terms:

-   Total *employment* `r change(labour_force, filter_with = filter_list, ym = 'month')` people.
    -   Over the past year, employment in `r params$state` has `r growth(labour_force, filter_with = filter_list,  ym = 'year')`.
-   Total *unemployment* `r filter_list$indicator = "Unemployed total"; change(labour_force, filter_with = filter_list, ym = 'month')` people, and the *unemployment rate* `r filter_list$indicator = "Unemployment rate";  change(labour_force, filter_with = filter_list, ym = 'month')`.
    -   Over the past year, unemployment in `r params$state` has `r filter_list$indicator = "Unemployed total"; growth(labour_force, filter_with = filter_list, ym = 'year')`, and the unemployment rate has `r filter_list$indicator = "Unemployment rate"; growth(labour_force, filter_with = filter_list, ym = 'year')`.
-   Total *underutilisation* has `r filter_list$indicator = 'Underutilised total'; change(labour_force, filter_with = filter_list, ym = 'month')` people, and the *underutilisation rate* `r filter_list$indicator = "Underutilisation rate"` `r change(labour_force, filter_with = filter_list, ym = 'month')`.
    -   Over the past year, total underutilisation has `r filter_list$indicator = "Underutilised total"; growth(labour_force, filter_with = filter_list, ym = 'year')` and the underutilisation rate has `r filter_list$indicator = "Underutilisation rate"; growth(labour_force, filter_with = filter_list, ym = 'year')`.

**Nationally** in `r release(labour_force, 'month')`:

-   Total *employment* `r filter_list = list(indicator = "Employed total", sex = "Persons", age = "Total (age)", state = "Australia", series_type = params$series_type); change(labour_force, filter_with = filter_list, ym = 'month')` people.
    -   Over the past year, employment in Australia has `r growth(labour_force, filter_with = filter_list, ym = 'year')`.
-   Total *unemployment* `r filter_list$indicator = "Unemployed total"; change(labour_force, filter_with = filter_list, ym = 'month')` people and the *unemployment rate* `r filter_list$indicator = "Unemployment rate"` `r change(labour_force,  filter_with = filter_list, ym = 'month')`
    -   Over the past year, unemployment in Australia has `r filter_list$indicator = "Unemployed total"; growth(labour_force, filter_with = filter_list,  ym = 'year')` and the *unemployment rate* has `r filter_list$indicator = "Unemployment rate"; growth(labour_force, filter_with = filter_list, ym = 'year')`.
-   Total *underutilisation* `r filter_list$indicator = 'Underutilised total'; change(labour_force, filter_with = filter_list, ym = 'month')` people, and the *underutilisation rate* `r filter_list$indicator = "Underutilisation rate"; change(labour_force, filter_with = filter_list,  ym = "month")`.
    -   Over the past year, total underutilisation has `r filter_list$indicator = "Underutilised total"` `r growth(labour_force, filter_with = filter_list,  ym = 'year')` and the underutilisation rate has `r filter_list$indicator = "Underutilisation rate"` `r growth(labour_force, filter_with = filter_list,  ym = 'year')`.

\clearpage
\section{Employment}

`r {filter_list = list(indicator = "Employed total", state = params$state, sex = "Persons", age = "Total (age)", series_type = params$series_type)}`

-   In `r release(labour_force, 'month')`, `r str_to_lower(params$series_type)` employment in `r params$state` `r change(labour_force, filter_with = filter_list,  ym = 'month')` people.
-   Compared with `r release(labour_force, 'month')` last year, total employment in `r params$state` has `r growth(labour_force, filter_with = filter_list,  ym = 'year')`.
-   Nationally, employment in `r params$state; filter_list$state = "Australia"; release(labour_force, 'month'); change(labour_force, filter_with = filter_list,  ym = 'month')` people.

```{r}
#| label: employment

filter_list$state = params$state
abs_plot(labour_force, over = filter_list)
```

\subsection{Full and Part Time Employment}

`r {filter_list = list(state = params$state, indicator = "Employed full-time", sex = "Persons", series_type = params$series_type)}`

-   Full-time employment in `r params$state` `r change(labour_force,  filter_with = filter_list,  ym = 'month')` in `r release(labour_force, 'month')`, and has `r growth(labour_force, filter_with = filter_list,  ym = 'year')` over the year. `r {filter_list = list(indicator = "Employed part-time", series_type = "Original", state = params$state)}`
-   Part-time employment in `r params$state`, in original terms `r change(labour_force, filter_with = filter_list,  ym = 'month')` in `r release(labour_force, 'month')`. Part-time employment has `r growth(labour_force, filter_with = filter_list,  ym = 'year')` over the last year.

```{r}
#| label: full-part-time-employment

abs_plot(labour_force, 
         over = list(indicator = c("Employed part-time", "Employed full-time"),
                  state = params$state), 
         years = params$years, 
         compare_aus = FALSE) 

```

\clearpage
\section{Unemployment}

`r {filter_list = list(indicator = "Unemployed total", state = params$state, series_type = params$series_type)}`

-   In `r release(labour_force, 'month')`, the total number of unemployed people in `r params$state` `r change(labour_force, filter_with = filter_list,  ym = 'month')`.

-   Compared with `r release(labour_force, 'month')` last year, unemployment in `r params$state` has `r growth(labour_force, filter_with = filter_list,  ym = 'year')` people. `r filter_list$state = "Australia"`

-   Nationally, total unemployment in `r release(labour_force, 'month')` `r change(labour_force, filter_with = filter_list,  ym = 'month')` people.

```{r}
#| label: unemployment-level

filter_list$state = params$state
abs_plot(labour_force, filter_list)
```

\subsection{Male and Female Unemployment Rates}

`r {filter_list = list(sex = "Males", indicator = "Unemployment rate", state = params$state)}`

-   In `r release(labour_force, 'month')`, the `r str_to_lower(params$series_type)` unemployment rate for males in `r params$state` `r change(labour_force, filter_with = filter_list,  ym = 'month')`.
-   Compared with `r release(labour_force, 'month')` last year male unemployment in `r params$state` has `r growth(labour_force, filter_with = filter_list,  ym = 'year')`.
-   In the same period, the `r str_to_lower(params$series_type)` unemployment rate for females in `r params$state; filter_list$sex = "Females"; params$state; change(labour_force, filter_with = filter_list,  ym = 'month')`.
-   Compared with `r release(labour_force, 'month')` last year female unemployment in `r params$state` has `r growth(labour_force, filter_with = filter_list,  ym = 'year')`.

```{r}
#| label: male-female-unemployment
filter_list$sex = c("Males", "Females")
abs_plot(labour_force, over = filter_list, compare_aus = F)

```

\clearpage
\section{Underemployment}

`r {filter_list = list(state = params$state, series_type = params$series_type, indicator = "Underemployment rate (proportion of labour force)")}`

-   The underemployment rate ``` r change(labour_force, filter_with = filter_list,  ym = 'month')``r filter_list$indicator = "Underemployed total" ```. The total number of people underemployed has `r change(labour_force, filter_with = filter_list,  ym = 'month')` people.

-   Over the past year, total underemployment has `r growth(labour_force, filter_with = filter_list,  ym = 'year')`.

-   Nationally, the underemployment rate in `r release(labour_force,   ym = 'month')` `r filter_list$state = "Australia"; filter_list$indicator = "Underemployment rate (proportion of labour force)"` `r change(labour_force, filter_with = filter_list,  ym = 'month')` and the total number of people underemployed `r filter_list$indicator = "Underemployed total"` `r change(labour_force, filter_with = filter_list, ym = 'month')`.

```{r}
#| label: fig-underemployment-rate
filter_list$state = params$state
filter_list$indicator = "Underemployment rate (proportion of labour force)"
abs_plot(labour_force, filter_list)
```

```{r}
#| label: fig-underemployment-number
filter_list$state = params$state
filter_list$indicator = "Underemployed total"
abs_plot(labour_force, filter_list)
```

\clearpage
\subsection{Hours Worked}

`r  {filter_list = list(state = params$state, series_type = params$series_type, indicator = "Monthly hours worked in all jobs")}`

-   The total number of hours worked in all jobs `r change(hours_worked, filter_with = filter_list,  ym = 'month')`.

-   Over the past year, the number of hours worked in `r params$state` has `r growth(hours_worked, filter_with = filter_list,  ym = 'year')`.

-   Nationally, `r filter_list$state = "Australia"` the total number of hours worked in `r release(hours_worked, ym = 'month')` `r change(hours_worked, filter_with = filter_list,  ym = 'month')`.

```{r}
#| label: fig-hours-worked
filter_list$state = params$state
abs_plot(hours_worked, filter_list)
```
