---
title: "Labour Market Brief  \n`r params$state`"
output: 
  pdf_document: 
    latex_engine: lualatex
    keep_tex: FALSE
fontsize: 11pt
geometry: margin = 1in, bottom = 1.5in
documentclass: article
classoption: a4paper
mainfont: Roboto

header-includes:
- \usepackage{fancyhdr}
- \usepackage{titling}
- \usepackage{comment}
- \usepackage{transparent}
- \fancypagestyle{first}{\fancyhf{}\setlength{\headheight}{120pt}\fancyhead[C]{\includegraphics[height=40mm]{img/FL_AITI_LOGO2}}\renewcommand{\footrulewidth}{0pt}\renewcommand{\headrulewidth}{0pt}}
- \fancypagestyle{common}{\fancyhf{}\setlength{\headheight}{15pt}\fancyfoot[C]{\vspace{10mm}\thepage}\fancyfoot[L]{{\transparent{0.75}\includegraphics[height = 20mm]{img/aiti_logo.png}}}\fancyfoot[R]{{\transparent{0.75}\includegraphics[height = 20mm]{`r paste0("img/", strayr::strayr(params$state),"_map.png")`}}}\fancyhead[L]{\textbf{Labour Market Brief}}\fancyhead[C]{}\fancyhead[R]{`r params$state`}\renewcommand{\footrulewidth}{1.0pt}}


params:
  series_type: Seasonally Adjusted
  state: South Australia
  years: 2017
  run: TRUE


---


```{r setup, include = FALSE, eval = TRUE}

#Required Packages
library(dplyr)
library(stringr)
library(tidyr)
library(ggplot2)
library(lubridate)
library(daitir)
library(reportabs)
library(scales)

#Need to create a trend part-time employment for the states
labour_force <- labour_force %>% 
  pivot_wider(names_from = indicator, 
    values_from = value) %>%
  mutate(`Share full-time` = `Employed full-time`/`Employed total`) %>%
  pivot_longer(cols = c(9:length(.)), names_to = 'indicator', values_to = 'value', values_drop_na = TRUE)

filter_list <- list(indicator = "Employed total", gender = "Persons", age = "Total (age)", state = params$state, series_type = params$series_type)

```

\section{Summary}
\thispagestyle{first}


The ABS Labour Force Survey data for **`r params$state`** released in `r release(labour_force, 'month', 1)` for `r release(labour_force, 'month')` showed that in **`r str_to_lower(params$series_type)`** terms:

+ Total *employment* `r change(labour_force, filter_with = filter_list, ym = 'month')` people. 
    + Over the past year, employment in `r params$state`  has `r growth(labour_force, filter_with = filter_list,  ym = 'year')`.
  
+ Total *unemployment* `r filter_list$indicator = "Unemployed total"; change(labour_force, filter_with = filter_list, ym = 'month')` people, and the *unemployment rate* `r filter_list$indicator = "Unemployment rate";  change(labour_force, filter_with = filter_list, ym = 'month')`. 
    + Over the past year, unemployment in `r params$state` has `r filter_list$indicator = "Unemployed total"; growth(labour_force, filter_with = filter_list, ym = 'year')`, and the unemployment rate has `r filter_list$indicator = "Unemployment rate"; growth(labour_force, filter_with = filter_list, ym = 'year')`. 

+ Total *underutilisation* has `r filter_list$indicator = 'Underutilised total'; change(labour_force, filter_with = filter_list, ym = 'month')` people, and the *underutilisation rate* `r filter_list$indicator = "Underutilisation rate"` `r change(labour_force, filter_with = filter_list, ym = 'month')`.
    + Over the past year, total underutilisation has `r filter_list$indicator = "Underutilised total"; growth(labour_force, filter_with = filter_list, ym = 'year')` and the underutilisation rate has `r filter_list$indicator = "Underutilisation rate"; growth(labour_force, filter_with = filter_list, ym = 'year')`. 

**Nationally** in `r release(labour_force, 'month')`:

+ Total *employment* `r filter_list = list(indicator = "Employed total", gender = "Persons", age = "Total (age)", state = "Australia", series_type = params$series_type); change(labour_force, filter_with = filter_list, ym = 'month')` people.
    + Over the past year, employment in Australia has `r growth(labour_force, filter_with = filter_list, ym = 'year')`. 
    
+ Total *unemployment* `r filter_list$indicator = "Unemployed total"; change(labour_force, filter_with = filter_list, ym = 'month')` people and the *unemployment rate* `r filter_list$indicator = "Unemployment rate"` `r change(labour_force,  filter_with = filter_list, ym = 'month')`
    + Over the past year, unemployment in Australia has `r filter_list$indicator = "Unemployed total"; growth(labour_force, filter_with = filter_list,  ym = 'year')` and the *unemployment rate* has `r filter_list$indicator = "Unemployment rate"; growth(labour_force, filter_with = filter_list, ym = 'year')`. 
    
+ Total *underutilisation* `r filter_list$indicator = 'Underutilised total'; change(labour_force, filter_with = filter_list, ym = 'month')` people, and the *underutilisation rate* `r filter_list$indicator = "Underutilisation rate"; change(labour_force, filter_with = filter_list,  ym = "month")`.
    + Over the past year, total underutilisation has `r filter_list$indicator = "Underutilised total"` `r growth(labour_force, filter_with = filter_list,  ym = 'year')` and the underutilisation rate has `r filter_list$indicator = "Underutilisation rate"` `r growth(labour_force, filter_with = filter_list,  ym = 'year')`.

\clearpage
\pagestyle{common}
\section{Employment}
`r {filter_list = list(indicator = "Employed total", state = params$state, gender = "Persons", age = "Total (age)", series_type = params$series_type)}`
    
+ In `r release(labour_force, 'month')`, `r str_to_lower(params$series_type)` employment in `r params$state` `r change(labour_force, filter_with = filter_list,  ym = 'month')` people. 
+ Compared with `r release(labour_force, 'month')` last year, total employment in `r params$state` has `r growth(labour_force, filter_with = filter_list,  ym = 'year')`. 
+ Nationally, employment in `r filter_list$state = "Australia"; release(labour_force, 'month'); change(labour_force, filter_with = filter_list,  ym = 'month')` people.


```{r employment, echo=FALSE,  fig.width = 8, fig.height = 4, units = 'cm'}
abs_plot("Employed total", params$state, params$years, series_types = params$series_type) +
  aes(linetype = state) +
  scale_linetype_manual(breaks = c(params$state, "Australia"), values = c(1,2))
```

\subsection{Full and Part Time Employment}
`r {filter_list = list(state = params$state, indicator = "Employed full-time", gender = "Persons", series_type = params$series_type)}`

+ Full-time employment in `r params$state` `r change(labour_force,  filter_with = filter_list,  ym = 'month')` in  `r release(labour_force, 'month')`, and has `r growth(labour_force, filter_with = filter_list,  ym = 'year')` over the year. `r {filter_list = list(indicator = "Employed part-time", series_type = "Original", state = params$state)}`
+ Part-time employment in `r params$state`, in original terms `r change(labour_force, filter_with = filter_list,  ym = 'month')` in `r release(labour_force, 'month')`. Part-time employment has `r growth(labour_force, filter_with = filter_list,  ym = 'year')` over the last year. 

``` {r full-part-time-employment, echo = F,  fig.height = 4, fig.width = 8, units = 'cm'}
plot_data <- labour_force %>%
  filter(indicator %in% c("Employed part-time", "Employed full-time"),
         gender == "Persons",
         state == params$state,
         age == "Total (age)",
    year >=  params$years) %>%
  group_by(indicator) %>%
  filter(series_type == ifelse(any(series_type == params$series_type), params$series_type, "Original")) %>%
  mutate(index = 100*value/value[1])

ggplot(plot_data, aes(x = date, y = index, colour = indicator, linetype = indicator)) +
  geom_line() +
  labs(
    x=NULL,
    subtitle = paste("Index (Base:", month(min(plot_data$date), abbr=F, label = T), year(min(plot_data$date)), "= 100)"),
    title = paste0("FULL AND PART-TIME EMPLOYMENT: ", strayr::strayr(params$state) %>% str_to_upper),
    caption = paste0("Source: 6202.0 - Labour Force, Australia, ", release(plot_data, ym = 'month'), " ", release(plot_data, ym = 'year'), " (Table 12, Derived)")
  ) +
  scale_x_date(date_breaks = reportabs:::date_breaks_format(params$years), labels = date_format("%b-%y")) +
  scale_colour_manual(breaks = c("Employed part-time", "Employed full-time"), values = c(aititheme::aiti_darkblue, aititheme::aiti_blue)) + 
  scale_linetype_manual(breaks = c("Employed part-time", "Employed full-time"), values = c("solid", "dashed")) + 
  aititheme::theme_aiti(legend = 'bottom') 
```

\clearpage
\section{Unemployment}
`r {filter_list = list(indicator = "Unemployed total", state = params$state, series_type = params$series_type)}`
+ In `r release(labour_force, 'month')`, the total number of unemployed people in `r params$state` `r change(labour_force, filter_with = filter_list,  ym = 'month')`.
+ Compared with `r release(labour_force, 'month')` last year, unemployment in `r params$state` has `r growth(labour_force, filter_with = filter_list,  ym = 'year')` people. `r filter_list$state = "Australia"`
+ Nationally, total unemployment in `r release(labour_force, 'month')` `r change(labour_force, filter_with = filter_list,  ym = 'month')` people.

```{r unemployment_level, echo=FALSE,  fig.height = 4, fig.width = 8, units = 'cm'}
abs_plot("Unemployed total", params$state, params$years,  series_types = params$series_type) +
  aes(linetype = state) +
  scale_linetype_manual(breaks = c(params$state, "Australia"), values = c(1,2))

```

\subsection{Unemployment Rate}
`r {filter_list = list(indicator = "Unemployment rate", state = params$state, series_type = params$series_type)}`

+ In `r release(labour_force, 'month')`, the `r str_to_lower(params$series_type)` unemployment rate in `r params$state` `r change(labour_force, filter_with = filter_list,  ym = 'month')` . 
+ Compared with `r release(labour_force, 'month')` last year, the unemployment rate in `r params$state` has `r growth(labour_force, filter_with = filter_list,  ym = 'year')`. 
+ Nationally, unemployment `r filter_list$state = "Australia"; release(labour_force, 'month'); change(labour_force, filter_with = filter_list,  ym = 'month')`

```{r unemployment-rate, echo=FALSE, fig.width = 8, fig.height = 4, units = 'cm'}
abs_plot("Unemployment rate", params$state, params$years, series_types = params$series_type) 
```

\subsection{Male and Female Unemployment Rates}
`r {filter_list = list(gender = "Males", indicator = "Unemployment rate", state = params$state)}`

+ In `r release(labour_force, 'month')`, the `r str_to_lower(params$series_type)` unemployment rate for males in `r params$state` `r change(labour_force, filter_with = filter_list,  ym = 'month')`. 
+ Compared with `r release(labour_force, 'month')` last year male unemployment in  `r params$state` has `r growth(labour_force, filter_with = filter_list,  ym = 'year')`.
+ In the same period, the `r str_to_lower(params$series_type)` unemployment rate for females in `r filter_list$gender = "Females"; params$state; change(labour_force, filter_with = filter_list,  ym = 'month')`. 
+ Compared with `r release(labour_force, 'month')` last year female unemployment in  `r params$state` has `r growth(labour_force, filter_with = filter_list,  ym = 'year')`. 


```{r male-female-unemployment, echo = FALSE,  fig.height = 4, fig.width = 8, units = 'cm'}

plot_data <- labour_force %>%
  filter(indicator == "Unemployment rate",
         gender != "Persons",
         state == params$state,
         age == "Total (age)",
         series_type == params$series_type, 
    year >= params$years)

ggplot(plot_data, aes(x = date, y = value, colour = gender, linetype = gender)) +
  geom_line() +
  labs(
    x=NULL,
    y=NULL,
    title = paste0("UNEMPLOYMENT RATES: MALE AND FEMALE: ", strayr::strayr(params$state) %>% str_to_upper),
    caption = str_c("Source: 6202.0 - Labour Force, Australia, ", release(plot_data, ym = 'month'), " ", release(plot_data, ym = 'year'), " (Table 12, ", params$series_type, ")")
  ) +
  scale_x_date(date_breaks = reportabs:::date_breaks_format(params$years), labels = date_format("%b-%y")) +
  scale_y_continuous(labels = percent_format(accuracy = 0.1, scale=1)) +
  scale_linetype_manual(breaks = c("Females", "Males"), values = c("solid", "dashed")) + 
  scale_colour_manual(breaks = c("Females", "Males"), values = c(aititheme::aiti_darkblue, aititheme::aiti_blue)) +
  aititheme::theme_aiti(legend = 'bottom')

```

\clearpage
\section{Underemployment}
`r {filter_list = list(state = params$state, series_type = params$series_type, indicator = "Underemployment rate (proportion of labour force)")}`
+ The underemployment rate `r change(labour_force, filter_with = filter_list,  ym = 'month')` `r filter_list$indicator = "Underemployed total"`. The total number of people underemployed has `r change(labour_force, filter_with = filter_list,  ym = 'month')` people.
+ Over the past year, total underemployment has `r growth(labour_force, filter_with = filter_list,  ym = 'year')`.
+ Nationally, the underemployment rate in `r release(labour_force,   ym = 'month')` `r filter_list$state = "Australia"; filter_list$indicator = "Underemployment rate (proportion of labour force)"` `r change(labour_force, filter_with = filter_list,  ym = 'month')` and the total number of people underemployed `r filter_list$indicator = "Underemployed total"` `r change(labour_force, filter_with = filter_list, ym = 'month')`

``` {r underemployment-rate, echo = F, fig.height = 4, fig.width = 8, units = 'cm'}
abs_plot("Underemployment rate (proportion of labour force)", params$state, params$years, series_type = params$series_type) +
  aes(linetype = state) +
  scale_linetype_manual(breaks = c(params$state, "Australia"), values = c(1,2))
```

``` {r underemployment-number, echo = F,  fig.height = 4, fig.width = 8, units = 'cm'}
abs_plot("Underemployed total", params$state, params$years, series_type = params$series_type) +
  aes(linetype = state) +
  scale_linetype_manual(breaks = c(params$state, "Australia"), values = c(1,2))
```

\clearpage
\section{Underutilisation}
`r {filter_list = list(state = params$state, series_type = params$series_type, indicator = "Underutilisation rate")}`
+ The underutilisation rate `r change(labour_force, filter_with = filter_list,  ym = 'month')`. `r filter_list$indicator = "Underutilised total"` Over the past month, the number of underutilised people has `r change(labour_force, filter_with = filter_list,  ym = 'month')`. 
+ Over the past year, underutilisation in `r params$state` has `r growth(labour_force, filter_with = filter_list,  ym = 'year')`.
+ Nationally `r filter_list$indicator = "Underutilisation rate"; filter_list$state = "Australia"` the underutilisation rate in `r release(labour_force, ym = 'month')` `r change(labour_force,  filter_with = filter_list, ym = 'month')`, and the total number of underutilised people `r filter_list$indicator = "Underutilised total"` `r change(labour_force, filter_with = filter_list,  ym = 'month')`. 

``` {r underutilisation-rate, echo = F, fig.height = 4, fig.width = 8, units = 'cm'}
abs_plot("Underutilisation rate", params$state, params$years, series_type = params$series_type) +
  aes(linetype = state) +
  scale_linetype_manual(breaks = c(params$state, "Australia"), values = c(1,2))
```

``` {r underutilisation-number, echo = F,  fig.height = 4, fig.width = 8, units = 'cm'}
abs_plot("Underutilised total", params$state, params$years, series_type = params$series_type) +
  aes(linetype = state) +
  scale_linetype_manual(breaks = c(params$state, "Australia"), values = c(1,2))
```
`r if(!params$run) {"\\begin{comment}"}`
\clearpage
\subsection{Hours Worked}
`r {if(params$run) {filter_list = list(state = params$state, series_type = params$series_type, indicator = "Monthly hours worked in all jobs")}}`
+ The total number of hours worked in all jobs `r change(labour_force, filter_with = filter_list,  ym = 'month')`. 
+ Over the past year, the number of hours worked in `r params$state` has `r growth(labour_force, filter_with = filter_list,  ym = 'year')`. 
+ Nationally, `r filter_list$state = "Australia"` the total number of hours worked in `r release(labour_force, ym = 'month')` `r change(labour_force, filter_with = filter_list,  ym = 'month')`.
`r if(!params$run) {"\\end{comment}"}`

```{r hours-worked, echo = F, fig.height = 4, fig.width = 8, units = 'cm', eval = params$run}
abs_plot("Monthly hours worked in all jobs", params$state, params$years, series_type = params$series_type) +
  aes(linetype = state) +
  scale_linetype_manual(breaks = c(params$state, "Australia"), values = c(1,2))
```

