---
title: "Labour Market Brief  \n`r params$state`"
output: 
  pdf_document: 
    latex_engine: lualatex
    keep_tex: FALSE
fontsize: 10pt
geometry: margin = 1in, bottom = 1.5in
documentclass: article
classoption: a4paper
mainfont: Roboto
header-includes:
- \usepackage{fancyhdr}
- \usepackage{titling}
- \usepackage{comment}
- \usepackage{transparent}
- \usepackage{geometry}
- \fancypagestyle{first}{\fancyhf{}\setlength{\headheight}{120pt}\fancyhead[C]{\includegraphics[height=40mm]{img/FL_AITI_LOGO2}}\renewcommand{\footrulewidth}{0pt}\renewcommand{\headrulewidth}{0pt}}
- \fancypagestyle{common}{\fancyhf{}\setlength{\headheight}{15pt}\fancyfoot[C]{\vspace{10mm}\thepage}\fancyfoot[L]{{\transparent{0.75}\includegraphics[height = 20mm]{img/aiti_logo.png}}}\fancyfoot[R]{{\transparent{0.75}\includegraphics[height = 20mm]{`r paste0("img/", strayr::clean_state(params$state),"_map.png")`}}}\fancyhead[L]{\textbf{Labour Market Brief}}\fancyhead[C]{}\fancyhead[R]{`r params$state`}\renewcommand{\footrulewidth}{1.0pt}}
params:
  series_type: Seasonally Adjusted
  state: South Australia
  years: 2017
  directory:
    value: NULL
  data: NULL
editor_options: 
  markdown: 
    wrap: sentence
---

```{r setup, include = F}
options(tinytex.verbose = TRUE)
#Required Packages
library(dplyr)
library(stringr)
library(tidyr)
library(ggplot2)
library(lubridate)
library(reportabs)
library(scales)
library(ragg)

ragg_png <- function(..., res = 300) {
  agg_png(..., res = res, units = "in")
}
knitr::opts_chunk$set(dev = "ragg_png")

#Create underutilised total indicator

data <- list("labour_force" = read_absdata("labour_force") |>
                 tidyr::pivot_wider(id_cols = c(date, sex, state, series_type, unit, age),
                                    names_from = indicator,
                                    values_from = value) |>
                 dplyr::mutate(`Underutilised total` = `Underemployed total` + `Unemployed total`) |>
                 tidyr::pivot_longer(cols = "Employed total":"Underutilised total",
                                     names_to = "indicator",
                                     values_to = "value",
                                     values_drop_na = TRUE),
               "hours_worked" = read_absdata("hours_worked"))

labour_force <- data$labour_force
hours_worked <- data$hours_worked
filter_list <- list(indicator = "Employed total", sex = "Persons", age = "Total (age)", state = params$state, series_type = params$series_type)

```

```{=tex}
\section{Summary}
\thispagestyle{first}
```
The ABS Labour Force Survey data for **`r params$state`** released in `r release(labour_force, 'month', 1)` for `r release(labour_force, 'month')` showed that in **`r str_to_lower(params$series_type)`** terms:

-   Total *employment* `r change(labour_force, filter_with = filter_list, ym = 'month')` people.
    -   Over the past year, employment in `r params$state` has `r growth(labour_force, filter_with = filter_list,  ym = 'year')`.
-   Total *unemployment* `r filter_list$indicator = "Unemployed total"; change(labour_force, filter_with = filter_list, ym = 'month')` people, and the *unemployment rate* `r filter_list$indicator = "Unemployment rate";  change(labour_force, filter_with = filter_list, ym = 'month')`.
    -   Over the past year, unemployment in `r params$state` has `r filter_list$indicator = "Unemployed total"; growth(labour_force, filter_with = filter_list, ym = 'year')`, and the unemployment rate has `r filter_list$indicator = "Unemployment rate"; growth(labour_force, filter_with = filter_list, ym = 'year')`.
-   Total *underutilisation* has `r filter_list$indicator = 'Underutilised total'; change(labour_force, filter_with = filter_list, ym = 'month')` people, and the *underutilisation rate* `r filter_list$indicator = "Underutilisation rate"` `r change(labour_force, filter_with = filter_list, ym = 'month')`.
    -   Over the past year, total underutilisation has `r filter_list$indicator = "Underutilised total"; growth(labour_force, filter_with = filter_list, ym = 'year')` and the underutilisation rate has `r filter_list$indicator = "Underutilisation rate"; growth(labour_force, filter_with = filter_list, ym = 'year')`.

**Nationally** in `r release(labour_force, 'month')`:

-   Total *employment* `r filter_list = list(indicator = "Employed total", sex = "Persons", age = "Total (age)", state = "Australia", series_type = params$series_type); change(labour_force, filter_with = filter_list, ym = 'month')` people.
    -   Over the past year, employment in Australia has `r growth(labour_force, filter_with = filter_list, ym = 'year')`.
-   Total *unemployment* `r filter_list$indicator = "Unemployed total"; change(labour_force, filter_with = filter_list, ym = 'month')` people and the *unemployment rate* `r filter_list$indicator = "Unemployment rate"` `r change(labour_force,  filter_with = filter_list, ym = 'month')`
    -   Over the past year, unemployment in Australia has `r filter_list$indicator = "Unemployed total"; growth(labour_force, filter_with = filter_list,  ym = 'year')` and the *unemployment rate* has `r filter_list$indicator = "Unemployment rate"; growth(labour_force, filter_with = filter_list, ym = 'year')`.
-   Total *underutilisation* `r filter_list$indicator = 'Underutilised total'; change(labour_force, filter_with = filter_list, ym = 'month')` people, and the *underutilisation rate* `r filter_list$indicator = "Underutilisation rate"; change(labour_force, filter_with = filter_list,  ym = "month")`.
    -   Over the past year, total underutilisation has `r filter_list$indicator = "Underutilised total"` `r growth(labour_force, filter_with = filter_list,  ym = 'year')` and the underutilisation rate has `r filter_list$indicator = "Underutilisation rate"` `r growth(labour_force, filter_with = filter_list,  ym = 'year')`.

```{=tex}
\clearpage
\pagestyle{common}
\section{Employment}
```
`r {filter_list = list(indicator = "Employed total", state = params$state, sex = "Persons", age = "Total (age)", series_type = params$series_type)}`

-   In `r release(labour_force, 'month')`, `r str_to_lower(params$series_type)` employment in `r params$state` `r change(labour_force, filter_with = filter_list,  ym = 'month')` people.
-   Compared with `r release(labour_force, 'month')` last year, total employment in `r params$state` has `r growth(labour_force, filter_with = filter_list,  ym = 'year')`.
-   Nationally, employment in `r params$state; filter_list$state = "Australia"; release(labour_force, 'month'); change(labour_force, filter_with = filter_list,  ym = 'month')` people.

```{r employment}
abs_plot(labour_force, filter_list) 
```

\subsection{Full and Part Time Employment}

`r {filter_list = list(state = params$state, indicator = "Employed full-time", sex = "Persons", series_type = params$series_type)}`

-   Full-time employment in `r params$state` `r change(labour_force,  filter_with = filter_list,  ym = 'month')` in `r release(labour_force, 'month')`, and has `r growth(labour_force, filter_with = filter_list,  ym = 'year')` over the year. `r {filter_list = list(indicator = "Employed part-time", series_type = "Original", state = params$state)}`
-   Part-time employment in `r params$state`, in original terms `r change(labour_force, filter_with = filter_list,  ym = 'month')` in `r release(labour_force, 'month')`. Part-time employment has `r growth(labour_force, filter_with = filter_list,  ym = 'year')` over the last year.

```{r full-part-time-employment}

abs_plot(labour_force, 
         v = list(indicator = c("Employed part-time", "Employed full-time"),
                  state = params$state), 
         years = params$years, 
         compare_aus = FALSE) +
  ggtitle("")

```

```{=tex}
\clearpage
\section{Unemployment}
```
`r {filter_list = list(indicator = "Unemployed total", state = params$state, series_type = params$series_type)}` + In `r release(labour_force, 'month')`, the total number of unemployed people in `r params$state` `r change(labour_force, filter_with = filter_list,  ym = 'month')`.
+ Compared with `r release(labour_force, 'month')` last year, unemployment in `r params$state` has `r growth(labour_force, filter_with = filter_list,  ym = 'year')` people.
`r filter_list$state = "Australia"` + Nationally, total unemployment in `r release(labour_force, 'month')` `r change(labour_force, filter_with = filter_list,  ym = 'month')` people.

```{r unemployment_level, echo=FALSE,  fig.height = 4, fig.width = 8, units = 'cm'}
abs_plot(labour_force, filter_list)
```

\subsection{Unemployment Rate}

`r {filter_list = list(indicator = "Unemployment rate", state = params$state, series_type = params$series_type)}`

-   In `r release(labour_force, 'month')`, the `r str_to_lower(params$series_type)` unemployment rate in `r params$state` `r change(labour_force, filter_with = filter_list,  ym = 'month')`.
-   Compared with `r release(labour_force, 'month')` last year, the unemployment rate in `r params$state` has `r growth(labour_force, filter_with = filter_list,  ym = 'year')`.
-   Nationally, unemployment `r filter_list$state = "Australia"; release(labour_force, 'month'); change(labour_force, filter_with = filter_list,  ym = 'month')`

```{r unemployment-rate}
abs_plot(labour_force, filter_list)
```

\subsection{Male and Female Unemployment Rates}

`r {filter_list = list(sex = "Males", indicator = "Unemployment rate", state = params$state)}`

-   In `r release(labour_force, 'month')`, the `r str_to_lower(params$series_type)` unemployment rate for males in `r params$state` `r change(labour_force, filter_with = filter_list,  ym = 'month')`.
-   Compared with `r release(labour_force, 'month')` last year male unemployment in `r params$state` has `r growth(labour_force, filter_with = filter_list,  ym = 'year')`.
-   In the same period, the `r str_to_lower(params$series_type)` unemployment rate for females in `r params$state; filter_list$sex = "Females"; params$state; change(labour_force, filter_with = filter_list,  ym = 'month')`.
-   Compared with `r release(labour_force, 'month')` last year female unemployment in `r params$state` has `r growth(labour_force, filter_with = filter_list,  ym = 'year')`.

```{r male-female-unemployment}


plot_data <- labour_force %>%
  filter(indicator == "Unemployment rate",
         sex != "Persons",
         state == params$state,
         age == "Total (age)",
         series_type == params$series_type, 
         year(date) >= params$years)

ggplot(plot_data, aes(x = date, y = value, colour = sex, linetype = sex)) +
  geom_line() +
  labs(
    x = NULL,
    y = NULL,
    title = paste0("UNEMPLOYMENT RATES: MALE AND FEMALE: ", strayr::clean_state(params$state) |>  str_to_upper),
    caption = paste0("Source: 6202.0 - Labour Force, Australia, ", release(plot_data, ym = 'month'), " ", release(plot_data, ym = 'year'), " (Table 12, ", params$series_type, ")")
  ) +
  scale_x_date(breaks = scales::pretty_breaks(n = min(length(unique(plot_data$month)), 6)), labels = date_format("%b-%y")) +
  scale_y_continuous(labels = percent_format(accuracy = 0.1, scale = 1)) +
  scale_linetype_manual(breaks = c("Females", "Males"), values = c("solid", "dashed")) + 
  aititheme::scale_colour_aiti() +
  aititheme::theme_aiti()

```

```{=tex}
\clearpage
\section{Participation Rate}
```
`r {filter_list = list(state = params$state, series_type = params$series_type, indicator = "Participation rate")}` + The participation rate `r change(labour_force, filter_with = filter_list, ym = "month")`.
+ Over the past year, the participation rate has `r growth(labour_force, filter_with = filter_list, ym = "year")`.
+ Nationally, the participation rate in `r release(labour_force, ym = "month")` `r filter_list$state = "Australia"` `r change(labour_force, filter_with = filter_list, ym = "month")`.

```{r participation-rate, echo = F, fig.height = 4, fig.width = 8, units = "cm"}
abs_plot(labour_force, filter_list)
```

```{=tex}
\clearpage
\section{Underemployment}
```
`r {filter_list = list(state = params$state, series_type = params$series_type, indicator = "Underemployment rate (proportion of labour force)")}` + The underemployment rate ``` r change(labour_force, filter_with = filter_list,  ym = 'month')``r filter_list$indicator = "Underemployed total" ```. The total number of people underemployed has `r change(labour_force, filter_with = filter_list,  ym = 'month')` people. + Over the past year, total underemployment has `r growth(labour_force, filter_with = filter_list,  ym = 'year')`. + Nationally, the underemployment rate in `r release(labour_force,   ym = 'month')` `r filter_list$state = "Australia"; filter_list$indicator = "Underemployment rate (proportion of labour force)"` `r change(labour_force, filter_with = filter_list,  ym = 'month')` and the total number of people underemployed `r filter_list$indicator = "Underemployed total"` `r change(labour_force, filter_with = filter_list, ym = 'month')`.

```{r underemployment-rate, echo = F, fig.height = 4, fig.width = 8, units = 'cm'}
abs_plot(labour_force, filter_list)
```

```{r underemployment-number, echo = F,  fig.height = 4, fig.width = 8, units = 'cm'}
abs_plot(labour_force, filter_list)
```

```{=tex}
\clearpage
\section{Underutilisation}
```
`r {filter_list = list(state = params$state, series_type = params$series_type, indicator = "Underutilisation rate")}` + The underutilisation rate `r change(labour_force, filter_with = filter_list,  ym = 'month')`.
`r filter_list$indicator = "Underutilised total"` Over the past month, the number of underutilised people has `r change(labour_force, filter_with = filter_list,  ym = 'month')`.
+ Over the past year, underutilisation in `r params$state` has `r growth(labour_force, filter_with = filter_list,  ym = 'year')`.
+ Nationally `r filter_list$indicator = "Underutilisation rate"; filter_list$state = "Australia"` the underutilisation rate in `r release(labour_force, ym = 'month')` `r change(labour_force,  filter_with = filter_list, ym = 'month')`, and the total number of underutilised people `r filter_list$indicator = "Underutilised total"` `r change(labour_force, filter_with = filter_list,  ym = 'month')`.

```{r underutilisation-rate, echo = F, fig.height = 4, fig.width = 8, units = 'cm'}
abs_plot(labour_force, filter_list)
```

```{r underutilisation-number, echo = F,  fig.height = 4, fig.width = 8, units = 'cm'}
abs_plot(labour_force, filter_list)
```

```{=tex}
\clearpage
\subsection{Hours Worked}
```
`r  {filter_list = list(state = params$state, series_type = params$series_type, indicator = "Monthly hours worked in all jobs")}` + The total number of hours worked in all jobs `r change(hours_worked, filter_with = filter_list,  ym = 'month')`.
+ Over the past year, the number of hours worked in `r params$state` has `r growth(hours_worked, filter_with = filter_list,  ym = 'year')`.
+ Nationally, `r filter_list$state = "Australia"` the total number of hours worked in `r release(hours_worked, ym = 'month')` `r change(hours_worked, filter_with = filter_list,  ym = 'month')`.

```{r hours-worked, echo = F, fig.height = 4, fig.width = 8, units = 'cm'}
abs_plot(hours_worked, filter_list)
```
